---
title: "Option B+"
author: "Xinyi Yan"
date: "4/21/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

load pacakges and data
```{r setup, include=FALSE}
library(lubridate)
library(haven)
library(dplyr)
library(table1)
library(sjmisc)
library(magrittr)
library(data.table)
med=read_dta("/Users/xinyiyan/Desktop/Option B+/med_abs_labeled_appended_01Apr2020.dta")

```


diagnose before/during/after pregnancy
if moms' date of confirm postive unknown in blue card, check if the date is available in green card, we define the date of postive comfirmation as first visit date for the following analysis.
```{r}
bcna<-subset(med,is.na(med_gpconfpos_bc_dateconfirmpos)|med$med_gpconfpos_bc_dateconfirmpos == as.Date('2020-01-01')) 
bcnov1_1<-bcna
bcnov1_1$med_gphivdiag_gc_dateconfirmpos[bcnov1_1$med_gphivdiag_gc_dateconfirmpos==as.Date('2020-01-01')]<-NA 

g4<-subset(bcnov1_1,is.na(bcnov1_1$med_gphivdiag_gc_dateconfirmpos))
g4$firv<-NA
g4$firstvisit<-NA #first visit date missing in both green and blue card

preg2<-interval(as.Date(bcnov1_1$med_gpinfdob_infdob)-280,as.Date(bcnov1_1$med_gpinfdob_infdob)) #create a pregnancy window

#green card
bcnov1_1$firstvisit<-ifelse(as.Date(bcnov1_1$med_gphivdiag_gc_dateconfirmpos) %within% preg2,1,0) #define 1st visit during pregnancy as 1, not during pregnacy as 0 

gcvisit1<-subset(bcnov1_1,bcnov1_1$firstvisit==0) 
gcvisit1$firv<-ifelse(as.Date(gcvisit1$med_gphivdiag_gc_dateconfirmpos)<as.Date(gcvisit1$med_gpinfdob_infdob-280),0,2) #define 1st visit before pregnancy as 0, after as 2

gcvisit2<-subset(bcnov1_1,bcnov1_1$firstvisit==1)
gcvisit2$firv<-gcvisit2$firstvisit

g3<-rbind(gcvisit1,gcvisit2)


#blue card
bcvisit1<-subset(med,!is.na(med$med_gpconfpos_bc_dateconfirmpos)&med$med_gpconfpos_bc_dateconfirmpos!=as.Date('2020-01-01'))
preg3<-interval(as.Date(bcvisit1$med_gpinfdob_infdob)-280,as.Date(bcvisit1$med_gpinfdob_infdob))
bcvisit1$firstvisit<-ifelse(as.Date(bcvisit1$med_gpconfpos_bc_dateconfirmpos) %within% preg3,1,0)
bcvisit2<-subset(bcvisit1,bcvisit1$firstvisit==0)
bcvisit2$firv<-ifelse(as.Date(bcvisit2$med_gpconfpos_bc_dateconfirmpos)<as.Date(bcvisit2$med_gpinfdob_infdob)-280,0,2)
bcvisit3<-subset(bcvisit1,bcvisit1$firstvisit==1)
bcvisit3$firv<-bcvisit3$firstvisit
g2<-rbind(bcvisit2,bcvisit3)




#medfirv$firv: 0=before pregnancy/1=during/2=after
#medfirv$firstvisit: 1=during
medfirv<-rbind(g2,g3,g4)
table(medfirv$firv)

medfirv$firv<-factor(as_factor(medfirv$firv),levels = c('Known Postives','Newly Diagnosed(during pregnancy','Newly Diagnosed(after pregnancy'))


```


Table1: Baseline characteritics
```{r,eval=TRUE}

medfirv$diag[as_factor(medfirv$firv)==0]<-'Known Postive'
medfirv$diag[as_factor(medfirv$firv)==1]<-'Newly diagnosed(during pregnancy)'
medfirv$diag[as_factor(medfirv$firv)==2]<-'Newly diagnosed(after pregnancy)'


#maternal age
medfirv$age[medfirv$med_mage==99|medfirv$med_mage==999]<-NA

medfirv$age<-cut(medfirv$med_mage,breaks=c(0,10,19,24,34,Inf),labels=c("< 10yr","10-19yr","20-24yr","25-43yr",">35yr"),
  right = TRUE)
label(medfirv$age)<-'Maternal Age'


#Partner tested for HIV
medfirv$med_bc_ptnrtest[medfirv$med_bc_ptnrtest==88]<-NA
medfirv$pt<-factor(as.factor(medfirv$med_bc_ptnrtest),labels = c('no','yes'))
label(medfirv$pt)<-'Partner tested for HIV'

#disclosed
medfirv$med_bc_ptnrdisc[medfirv$med_bc_ptnrdisc==88]<-NA
medfirv$disclosed<-factor(as.factor(medfirv$med_bc_ptnrdisc),labels = c('no','yes'))
label(medfirv$disclosed)<-'Disclosed'


#time of ART, create ART initiation window
medfirv$med_gpintart_bc_dateintart[medfirv$med_gpintart_bc_dateintart==as.Date('2020-01-01')]<-NA
mn<-subset(medfirv,is.na(medfirv$med_gpintart_bc_dateintart))
mn$med_gpintartgc_gc_dateintart[mn$med_gpintartgc_gc_dateintart==as.Date('2020-01-01')]<-NA
mn$artdate<-mn$med_gpintartgc_gc_dateintart
mm<-subset(medfirv,!is.na(medfirv$med_gpintart_bc_dateintart))
mm$artdate<-mm$med_gpintart_bc_dateintart
medart<-rbind(mm,mn)
firsttri<-interval((as.Date(medart$med_gpinfdob_infdob))-280,as.Date(medart$med_gpinfdob_infdob)-190)
sectri<-interval((as.Date(medart$med_gpinfdob_infdob))-190,as.Date(medart$med_gpinfdob_infdob)-100)
thtri<-interval((as.Date(medart$med_gpinfdob_infdob))-100,as.Date(medart$med_gpinfdob_infdob)-10)
delivery<-interval((as.Date(medart$med_gpinfdob_infdob))-10,as.Date(medart$med_gpinfdob_infdob)+10)
post<-interval((as.Date(medart$med_gpinfdob_infdob))+10,as.Date(medart$med_gpinfdob_infdob)+Inf)
before<-interval((as.Date(medart$med_gpinfdob_infdob))-Inf,as.Date(medart$med_gpinfdob_infdob)-280)
medart$art[medart$artdate %within%firsttri]<-1
medart$art[medart$artdate %within%sectri]<-2
medart$art[medart$artdate %within%thtri]<-3
medart$art[medart$artdate %within%delivery]<-4
medart$art[medart$artdate >= as.Date(medart$med_gpinfdob_infdob)+10]<-5
medart$art[medart$artdate <=as.Date(medart$med_gpinfdob_infdob)-280]<-6
medfirv$art<-medart$art
medfirv$art<-factor(medfirv$art, labels = c("1st trimmester","2nd trimester","3rd trimester","labour&delievery","postnatal","before pregnancy"))
table(medfirv$art)


#regimen
medfirv$med_hei_momartreg[medfirv$med_hei_momartreg==88]<-NA
medfirv$regimen<-factor(as.factor(medfirv$med_hei_momartreg),labels=c('None','sdNVP Only','Interrupted HAART','AZT+NVP+3TC','HAART','Other (specify)'))
label(medfirv$regimen)<-'Regimen started'
table(medfirv$med_hei_momartreg)
                        
#mode of delivery
medfirv$med_hei_delivmethod[medfirv$med_hei_delivmethod==88]<-NA
medfirv$mode<-factor(as.factor(medfirv$med_hei_delivmethod),labels=c('SVD','C-section'))
label(medfirv$mode)<-'Mode of delivery'


#place of delivery
medfirv$med_hei_delivplace[medfirv$med_hei_delivplace==88]<-NA
medfirv$place<-factor(as.factor(medfirv$med_hei_delivplace),labels=c('facility','home'))
label(medfirv$place)<-'Place of delivery'

#status of mother
medfirv$med_hei_momalive[medfirv$med_hei_momalive==88]<-NA
medfirv$status<-factor(as.factor(medfirv$med_hei_momalive),labels=c('Dead','Alive'))
label(medfirv$status)<-'Status of mother' 

table1(~age+pt+disclosed+art+regimen+mode+place+status|diag,data=medfirv)

```


Table 2:Contrast between retained in care and drop off group
```{r}
#exclude moms that do not have HEI card
med1<-med
med1$hei<-ifelse(med1$med_forms1==3|med1$med_forms2==3|med1$med_forms3==3|med1$med_forms4==3,1,0) 
table(med1$hei) #6035 HEI
sum(is.na(med1$hei)) ##55 NA
infhei<-subset(med1,med1$hei==1) 


#combine med_heifinaloutcome and med_infoutcome
#use med_infoutcome to fill in the missing in med_heifinaloutcome
infhei$med_heifinaloutcome[infhei$med_heifinaloutcome==88|infhei$med_heifinaloutcome==0|infhei$med_heifinaloutcome==9]<-NA
inf11<-subset(infhei,is.na(infhei$med_heifinaloutcome))
inf11$outcome<-inf11$med_infoutcome
inf11$outcome[inf11$outcome==0|inf11$outcome==88]<-NA
inf22<-subset(infhei,!is.na(infhei$med_heifinaloutcome))
inf22$outcome<-inf22$med_heifinaloutcome
t2dat<-rbind(inf11,inf22)
t2datt<-subset(t2dat,!is.na(t2dat$outcome))

#retained in care:discharged, transfer out, referred to ccc, dead
#drop off: LTFU
t2datt$care<-(ifelse(t2datt$outcome==1|t2datt$outcome==2|t2datt$outcome==3|t2datt$outcome==5,"Retained in care",'Drop off'))

infdrop<-subset(t2dat,is.na(t2dat$outcome))
infdrop$care<-'Drop off' #define missing in infantoutcome as 'drop off'

infdrop$hei<-99



infdat<-rbind(t2datt,infdrop)

table(infdat$care)

inff1<-subset(infdat,infdat$care=='Drop off')
inff2<-subset(infdat,infdat$care=='Retained in care')

#create the drop off window
inff1<-mutate(inff1,id=rownames(inff1))
df2<-data.frame(id=inff1$id,d_23=as.Date(inff1$hei_visdate23), d_22=as.Date(inff1$hei_visdate22),d_21=as.Date(inff1$hei_visdate21),d_20=as.Date(inff1$hei_visdate20),d_19=as.Date(inff1$hei_visdate19),d_18=as.Date(inff1$hei_visdate18),d_17=as.Date(inff1$hei_visdate17),d_16=as.Date(inff1$hei_visdate16),d_15=as.Date(inff1$hei_visdate15),d_14=as.Date(inff1$hei_visdate14),d_13=as.Date(inff1$hei_visdate13),d_12=as.Date(inff1$hei_visdate12),d_11=as.Date(inff1$hei_visdate11),d_10=as.Date(inff1$hei_visdate10),d_9=as.Date(inff1$hei_visdate9),d_8=as.Date(inff1$hei_visdate8),d_7=as.Date(inff1$hei_visdate7),d_6=as.Date(inff1$hei_visdate6),d_5=as.Date(inff1$hei_visdate5),d_4=as.Date(inff1$hei_visdate4),d_3=as.Date(inff1$hei_visdate3),d_2=as.Date(inff1$hei_visdate2),d_1=as.Date(inff1$hei_visdate1))


df2$last<-do.call(pmax, c(df2[-1],na.rm=TRUE)) #find the first and last visit date
df2$first<-do.call(pmin, c(df2[-1],na.rm=TRUE))

inff1$lastvisit<-df2$last
inff1$firstvisit<-df2$first

iw1<-interval(as.Date(inff1$med_gpinfdob_infdob),as.Date(inff1$med_gpinfdob_infdob)+120) #0-3 months
iw2<-interval(as.Date(inff1$med_gpinfdob_infdob)+120,as.Date(inff1$med_gpinfdob_infdob)+240) #4-7
iw3<-interval(as.Date(inff1$med_gpinfdob_infdob)+240,as.Date(inff1$med_gpinfdob_infdob)+420) #8-13
iw4<-interval(as.Date(inff1$med_gpinfdob_infdob)+420,as.Date(inff1$med_gpinfdob_infdob)+600) #14-19
iw5<-interval(as.Date(inff1$med_gpinfdob_infdob)+600,as.Date(inff1$med_gpinfdob_infdob)+780) #20-25


iw13<-interval(as.Date(inff1$med_gpinfdob_infdob),as.Date(inff1$med_gpinfdob_infdob)+210)
inff1$fa[inff1$firstvisit %within%iw13]<-1


inff1$window[inff1$lastvisit %within%iw1]<-1 #moms dropped off between 0-3 months are labled as 1
inff1$window[inff1$lastvisit %within%iw2]<-2
inff1$window[inff1$lastvisit %within%iw3]<-3
inff1$window[inff1$lastvisit %within%iw4]<-4
inff1$window[inff1$lastvisit %within%iw5]<-5
table(inff1$window)

inff2$window<-NA #those retained in care do not have drop off windown
inff2$lastvisit<-NA
inff2$firstvisit<-NA
inff2$fa<-NA
inff2<-mutate(inff2,id=rownames(inff2))

infdath<-rbind(inff1,inff2)


#make the table
#infant sex
infdath$sex1[infdath$med_hei_sex==2]<-'Female'
infdath$sex1[infdath$med_hei_sex==1]<-'Male'
infdath$sex1[infdath$med_hei_sex==88]<-NA
label(infdath$sex1)<-'Sex'

                   
#infant age
infdath$age<-cut(infdath$med_gpageenroll_hei_ageenrollwk,c(0,8,12,48,Inf),labels=c('< 8 weeks','8 weeks-6 months','6-12 months','> 12 months'))

infdath$agem<-cut(infdath$med_gpageenroll_hei_ageenrollmo,c(0,2,6,12,Inf),labels=c('< 8 weeks','8 weeks-6 months','6-12 months','> 12 months'))
infdath$age1[(infdath$med_gpageenroll_hei_ageenrollwk<8)]='< 8 weeks'

infdath$age1[infdath$med_gpageenroll_hei_ageenrollmo>=6&infdath$med_gpageenroll_hei_ageenrollmo<=12]='6-12 months'

infdath$age1[infdath$med_gpageenroll_hei_ageenrollmo>12]="> 12 months"

label(infdath$age1)<-'Infant Age at enrollment'

#entry point
infdath$med_hei_entrypoint[infdath$med_hei_entrypoint==88]<-NA
infdath$entry<-factor(infdath$med_hei_entrypoint,labels=c('Paediatric ward','OPD','Maternity','CCC','MCH/PMTCT','Other (specify)'))
label(infdath$entry)<-'Clinical Entry Point'


#CTX
infdath$ctx<-ifelse(infdath$hei_ctx1=='0'|infdath$hei_ctx1=='00'|infdath$hei_ctx1=='999'|infdath$hei_ctx1=="N"|infdath$hei_ctx1=='O'|is.na(infdath$hei_ctx1),"No","Yes")
label(infdath$ctx)<-'Received CTX Prophylaxis'

#infant prophylaxis
infdath$med_hei_arvs[infdath$med_hei_arvs==88]<-NA
infdath$prophylaxis<-factor(infdath$med_hei_arvs,labels = c( 'None','Sd NVP only','Sd NVP+AZT+3TC','NVP for 6 weeks (Mother on ART or No BF)','NVP during breastfeeding','Other (specify)'))
infdath$prophylaxis[infdath$med_hei_arvs==0]='None'
label(infdath$prophylaxis)<-'Infant PMTCT Prophylaxis'



#drop off window
infdath$window<-as.factor(infdath$window)
label(infdath$window)<-"drop off window"



#feed: EBM/Emf/NBF/No breastfeed/WBF/Dbf/Erf/NVP/RF 
table(infdath$hei_inffeed2)
infdath$feed[infdath$hei_inffeed1==0|infdath$hei_inffeed1=="1.5mls"|infdath$hei_inffeed1==47||infdath$hei_inffeed1==54||infdath$hei_inffeed1==999||infdath$hei_inffeed1==1.5||infdath$hei_inffeed1==2||infdath$hei_inffeed1=="2MLS"||infdath$hei_inffeed1==51||infdath$hei_inffeed1==64||infdath$hei_inffeed1=="Yes"||infdath$hei_inffeed1=="Y"|infdath$hei_inffeed1=="Exclusive"|infdath$hei_inffeed1=="EBR"|infdath$hei_inffeed1=="Ebg"|infdath$hei_inffeed1=="Dbf"|infdath$hei_inffeed1=="WBF"]<-NA

infdath$feed[infdath$hei_inffeed1=="E bf"|infdath$hei_inffeed1=="Ebf"|infdath$hei_inffeed1=="EBF exclusive breastfeeding"||infdath$hei_inffeed1=="Exclusive  breastfeeding"|infdath$hei_inffeed1=="Exclusive breasfeeding"|infdath$hei_inffeed1=="exclusive breast feeding"|infdath$hei_inffeed1=="Exclusive breastf eeding"|infdath$hei_inffeed1==" Exclusive breastfeeding"|infdath$hei_inffeed1=="Exclusive breastfeeding  (EBF)"|infdath$hei_inffeed1=="Exclusivebreastfeeding"|infdath$hei_inffeed1=="EB"|infdath$hei_inffeed1=="EBF"|infdath$hei_inffeed1=="Exclusive  Breastfeeding"|infdath$hei_inffeed1=="Exclusive breast feeding"|infdath$hei_inffeed1=="Exclusive breast breastfeeding"|infdath$hei_inffeed1=="Exclusive Breastfeeding"|infdath$hei_inffeed1=="Exclusive breastfeeding Breast"|infdath$hei_inffeed1=="Ebf"|infdath$hei_inffeed1=="=Ebf"]<-"Exclusive Breastfeeding"


infdat$feed[infdat$hei_inffeed1=="ERF"|infdat$hei_inffeed1=="Exclusive replacement feed"|infdat$hei_inffeed1=="Exclusive Replacement feeding"|infdat$hei_inffeed1=="Erf"|infdat$hei_inffeed1=="Exclusive replacement feeding"|infdat$hei_inffeed1=="Exclusive Replacement Feeding"|infdat$hei_inffeed1=="RF"]<-"Exlusive Replacement Feeding"

infdath$feed[infdath$hei_inffeed1=="Mbf"|infdath$hei_inffeed1=="MF"|infdath$hei_inffeed1=="Mixed feeding"|infdath$hei_inffeed1=="Breastfeeding  and Complementary feeding"|infdath$hei_inffeed1=="Mb"|infdath$hei_inffeed1=="Mf"|infdath$hei_inffeed1=="Mixed feed"|infdath$hei_inffeed1=="Bebf"|infdath$hei_inffeed1=='BF'|infdath$hei_inffeed1=="Breastfeed"|infdath$hei_inffeed1=="Breastfeeding  (BF )"|infdath$hei_inffeed1=="BBF"|infdath$hei_inffeed1=="Bf"|infdath$hei_inffeed1=="Breast feeding"|infdath$hei_inffeed1=="Breastfeeding"|infdath$hei_inffeed1=="Cf"|infdath$hei_inffeed1=="Complementary feeding"|infdath$hei_inffeed1=="CF"]<-"Mix Feeding"


infdath$feed[infdath$hei_inffeed1=="Not breast feeding"|infdath$hei_inffeed1=="NVP"|infdath$hei_inffeed1=="No breastfeeding"|infdath$hei_inffeed1=="No"|infdath$hei_inffeed1=="Nbf"|infdath$hei_inffeed1=="Not breastfeeding"|infdath$hei_inffeed1=="Not BF"|infdath$hei_inffeed1=="No breastfeed"|infdath$hei_inffeed1=="NBF"]<-"No Breastfeeding"


#final outcome
infdath$med_heifinaloutcome[infdath$med_heifinaloutcome==88]<-NA
infdath$finaloutcome<-factor(infdath$med_heifinaloutcome,labels = c('Discharged at 18 months',"Referred to CCC","Transferred out","Lost to Follow up","Dead"))

table(infdath$finaloutcome)



#infant hiv status
#if the 1st pcr test or repeat PCR (1) test or repeat PCR (2) test or Final Antibody test or other test is postive, than the infant is HIV postive; if only the infant only has antibody test, we need to take a look at the infant age, if infant is younger than 14 months, we need infant to have both 1st Antibody test and confirmatory PCR test postive. If the infant is older than 14 months, we only need 1st Antibody test to be postive
infdath$hivstatus[infdath$med_gp1pcr_hei_res1pcr==1|infdath$med_gp1reppcr_hei_res1reppcr==1|infdath$med_gp2reppcr_hei_res2reppcr==1|infdath$med_gpfinab_hei_resfinab==1|infdath$med_gpoth_hei_resoth==1]<-"postive"

infdath$noother<-ifelse(is.na(infdath$med_gp1pcr_hei_res1pcr)&is.na(infdath$med_gp1reppcr_hei_res1reppcr)&is.na(infdath$med_gp2reppcr_hei_res2reppcr)&is.na(infdath$med_gpfinab_hei_resfinab)&is.na(infdath$med_gpoth_hei_resoth),0,1) 

infdath$ab<-ifelse(infdath$noother==0&!is.na(infdath$med_gp1ab_hei_res1ab),1,0)

infdath$young<-ifelse(infdath$ab==1&as.Date(infdath$med_gp1ab_hei_date1ab)-as.Date(infdath$med_gpinfdob_infdob)<420,1,0)

infdath$hivstatus[infdath$young==1&infdath$med_gpconfpcr_hei_resconfpcr==1&infdath$med_gp1ab_hei_res1ab==1]<-"postive"
infdath$hivstatus[infdath$young==0&infdath$med_gp1ab_hei_res1ab==1]<-"postive"


infdath$hivstatus[infdath$med_gp1pcr_hei_res1pcr==0|infdath$med_gp1reppcr_hei_res1reppcr==0|infdath$med_gp2reppcr_hei_res2reppcr==0|infdath$med_gpfinab_hei_resfinab==0|infdath$med_gpoth_hei_resoth==0|infdath$med_gp1ab_hei_res1ab==0|infdath$med_gpconfpcr_hei_resconfpcr==0]<-"negative"


sum(is.na(infdat$hivstatus))
table(infdath$hivstatus)



infdatcheck<-subset(infdath,infdath$care=="Drop off" & infdath$hivstatus=="postive")
table(infdatcheck$finaloutcome)

infdathh<-infdath

infdathh$care[infdathh$care=="Drop off" & infdathh$hivstatus=="postive"]<-"Retained in care" # if infant's HIV status is postivem than he/she is moved to 'retained in care'
table(infdathh$care)
table1(~window+sex1+age1+entry+ctx+prophylaxis+feed+hivstatus+finaloutcome|care,data=infdathh)


#chi-square test
chisq.test(infdathh$care,infdathh$finaloutcome)
chisq.test(infdathh$care,infdathh$hivstatus)
chisq.test(infdathh$care,infdathh$feed)
chisq.test(infdathh$care,infdathh$prophylaxis)
chisq.test(infdathh$care,infdathh$ctx)
chisq.test(infdathh$care,infdathh$entry)
chisq.test(infdathh$care,infdathh$age1)
chisq.test(infdathh$care,infdathh$sex1)

```






Table 3 : Contrast between early and late drop off
```{r}
infdathh$window[infdathh$care=="Retained in care"]<-NA
infdathhh<-subset(infdathh,!is.na(infdathh$window))
infdathhh$drop[infdathhh$window==1|infdathhh$window==2|infdathhh$window==3]<-0 #early drop off
infdathhh$drop[infdathhh$window==4|infdathhh$window==5]<-1 #late drop off

table1(~sex1+age1+entry+ctx+prophylaxis+feed+hivstatus+finaloutcome|drop,data=infdathhh)


chisq.test(infdathhh$drop,infdathhh$sex1)$p.value
chisq.test(infdathhh$drop,infdathhh$age1)$p.value
chisq.test(infdathhh$drop,infdathhh$entry)$p.value
chisq.test(infdathhh$drop,infdathhh$ctx)$p.value
chisq.test(infdathhh$drop,infdathhh$prophylaxis)$p.value
chisq.test(infdathhh$drop,infdathhh$feed)$p.value




```


Poisson Regression

poi.1<-gee(aa ~ sex1+age1+entry+ctx+prophylaxis+feed, family = poisson,id = seq(1,nrow(infdathh)), data=infdathh)
summary(poi.1) #risk ratio

infdathh$aa<-ifelse(infdathh$care=="Drop off",1,0)
poi.2<-gee(drop~sex1+age1+entry+ctx+prophylaxis+feed+hivstatus,family = poisson,id = seq(1,nrow(infdathhh)), data=infdathhh)
summary(poi.2)


Poisson regression+baseline characteristics
```{r}
#Risk ratio of dropping off vs retained in care
infdathh$aa<-ifelse(infdathh$care=="Drop off",1,0)
poi1<-regress('rate',aa ~ as.factor(sex1)+as.factor(age1)+as.factor(entry)+as.factor(ctx)+as.factor(prophylaxis)+as.factor(feed)+as.factor(hivstatus)+as.factor(feed)*as.factor(age1)+as.factor(momage)+as.factor(art)+as.factor(regimen)+as.factor(place),data=infdathh) #adjust for possible interactions

print(round(poi1$transformed,4))




# risk ratio of late drop off vs early drop off
infdathhh$entry[infdathhh$entry=='Missing']<-NA
infdathhh$prophylaxis[infdathhh$prophylaxis=='Missing']<-NA

poi2<-regress('rate',drop ~ as.factor(sex1)+as.factor(age1)+as.factor(entry)+as.factor(ctx)+as.factor(prophylaxis)+as.factor(feed)+as.factor(feed)*as.factor(age1)+as.factor(momage)+as.factor(regimen)+as.factor(place),data=infdathhh)
print(round(poi2$transformed,4))





```


exclude babies older than 6 month
```{r}
#create an age at 1st visit feature
df3<-data.frame(id=infdath$id,d_23=as.Date(infdath$hei_visdate23), d_22=as.Date(inff1$hei_visdate22),d_21=as.Date(inff1$hei_visdate21),d_20=as.Date(inff1$hei_visdate20),d_19=as.Date(inff1$hei_visdate19),d_18=as.Date(inff1$hei_visdate18),d_17=as.Date(inff1$hei_visdate17),d_16=as.Date(inff1$hei_visdate16),d_15=as.Date(inff1$hei_visdate15),d_14=as.Date(inff1$hei_visdate14),d_13=as.Date(inff1$hei_visdate13),d_12=as.Date(inff1$hei_visdate12),d_11=as.Date(inff1$hei_visdate11),d_10=as.Date(inff1$hei_visdate10),d_9=as.Date(inff1$hei_visdate9),d_8=as.Date(inff1$hei_visdate8),d_7=as.Date(inff1$hei_visdate7),d_6=as.Date(inff1$hei_visdate6),d_5=as.Date(inff1$hei_visdate5),d_4=as.Date(inff1$hei_visdate4),d_3=as.Date(inff1$hei_visdate3),d_2=as.Date(inff1$hei_visdate2),d_1=as.Date(inff1$hei_visdate1))

infdath$fa[infdath$hei_visdate1%within%iw13]<-1

df2$last<-do.call(pmax, c(df2[-1],na.rm=TRUE))
df2$first<-do.call(pmin, c(df2[-1],na.rm=TRUE))

inff1$lastvisit<-df2$last
inff1$firstvisit<-df2$first


iw13<-interval(as.Date(infdath$med_gpinfdob_infdob),as.Date(infdath$med_gpinfdob_infdob)+210) 
inff1$fa[inff1$firstvisit %within%iw13]<-1  #younger than 6 months when 1st visit

table(infdath$fa)

infdathhhh<-subset(infdath,infdath$fa==1)
infdathhhh$window[infdathhhh$care=="Retained in care"]<-NA
infdathhhhh<-subset(infdathhhh,!is.na(infdathhhh$window))
infdathhhhh$drop[infdathhhhh$window==1|infdathhhhh$window==2|infdathhhhh$window==3]<-0
infdathhhhh$drop[infdathhhhh$window==4|infdathhhhh$window==5]<-1

table1(~sex1+age1+entry+ctx+prophylaxis+feed+hivstatus+finaloutcome|drop,data=infdathhhhh)

infdathhhhh$entry[infdathhhhh$entry=='Missing']<-NA
infdathhhhh$prophylaxis[infdathhhhh$prophylaxis=='Missing']<-NA
poi3<-regress('rate',drop ~ as.factor(sex1)+as.factor(age1)+as.factor(entry)+as.factor(ctx)+as.factor(prophylaxis)+as.factor(feed)+as.factor(feed)*as.factor(age1)+as.factor(momage)+as.factor(regimen)+as.factor(place),data=infdathhhhh)
print(round(poi3$transformed,4)) #same poission model as in model3
```


add baseline characteristics to tables
```{r}
#maternal age
infdath$momage[infdath$med_mage==99|infdath$med_mage==999]<-NA
infdath$momage<-cut(infdath$med_mage,breaks=c(0,24,34,Inf),labels=c("<=24yr","25-34yr",">35yr"),
  right = TRUE)
table(infdath$momage)
label(infdath$momage)<-'Maternal Age'

#time of ART
infdath$med_gpintart_bc_dateintart[infdath$med_gpintart_bc_dateintart==as.Date('2020-01-01')]<-NA
mn<-infdath[is.na(infdath$med_gpintart_bc_dateintart),]
mn$med_gpintartgc_gc_dateintart[mn$med_gpintartgc_gc_dateintart==as.Date('2020-01-01')]<-NA
mn$artdate<-mn$med_gpintartgc_gc_dateintart
mm<-infdath[!is.na(infdath$med_gpintart_bc_dateintart),]
mm$artdate<-mm$med_gpintart_bc_dateintart
medart<-rbind(mm,mn)
firsttri<-interval((as.Date(medart$med_gpinfdob_infdob))-280,as.Date(medart$med_gpinfdob_infdob)-190)
sectri<-interval((as.Date(medart$med_gpinfdob_infdob))-190,as.Date(medart$med_gpinfdob_infdob)-100)
thtri<-interval((as.Date(medart$med_gpinfdob_infdob))-100,as.Date(medart$med_gpinfdob_infdob)-10)
delivery<-interval((as.Date(medart$med_gpinfdob_infdob))-10,as.Date(medart$med_gpinfdob_infdob)+10)
post<-interval((as.Date(medart$med_gpinfdob_infdob))+10,as.Date(medart$med_gpinfdob_infdob)+Inf)
before<-interval((as.Date(medart$med_gpinfdob_infdob))-Inf,as.Date(medart$med_gpinfdob_infdob)-280)
medart$artinf[medart$artdate %within% firsttri]<-1
medart$artinf[medart$artdate %within%sectri]<-2
medart$artinf[medart$artdate %within%thtri]<-3
medart$artinf[medart$artdate %within%delivery]<-4
medart$artinf[medart$artdate >= as.Date(medart$med_gpinfdob_infdob)+10]<-5
medart$artinf[medart$artdate <=as.Date(medart$med_gpinfdob_infdob)-280]<-6
infdath$art<-medart$artinf
infdath$art<-factor(infdath$art, labels = c("1st trimmester","2nd trimester","3rd trimester","labour&delievery","postnatal","before pregnancy"))


#regimen
infdath$med_hei_momartreg[infdath$med_hei_momartreg==88]<-NA
infdath$regimen<-factor(as.factor(infdath$med_hei_momartreg),labels=c('None','sdNVP Only','Interrupted HAART','AZT+NVP+3TC','HAART','Other (specify)'))
label(infdath$regimen)<-'Regimen started'

#place of delivery
infdath$med_hei_delivplace[infdath$med_hei_delivplace==88]<-NA
infdath$place<-factor(as.factor(infdath$med_hei_delivplace),labels=c('facility','home'))
label(infdath$place)<-'Place of delivery'


infdathh<-infdath

infdathh$care[infdathh$care=="Drop off" & infdathh$hivstatus=="postive"]<-"Retained in care" # if infant's HIV status is postivem than he/she is moved to 'retained in care'

table1(~window+sex1+age1+entry+ctx+prophylaxis+feed+hivstatus+finaloutcome+momage+art+regimen+place|care,data=infdathh)


#retained in care vs drop off
table1(~window+sex1+age1+entry+ctx+prophylaxis+feed+hivstatus+finaloutcome+momage+art+regimen+place|care,data=infdathh)


#early vs late

infdathh$window[infdathh$care=="Retained in care"]<-NA
infdathhh<-subset(infdathh,!is.na(infdathh$window))
infdathhh$drop[infdathhh$window==1|infdathhh$window==2|infdathhh$window==3]<-'early' #early drop off
infdathhh$drop[infdathhh$window==4|infdathhh$window==5]<-'late' #late drop off
table1(~sex1+age1+entry+ctx+prophylaxis+feed+hivstatus+finaloutcome+momage+art+regimen+place|drop,data=infdathhh)



```

